---
title: "pudms"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pudms}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo = FALSE, include=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(pudms)
```

## Description

This package offers a streamlined analysis via PUlasso algorithm for learning sequence-function relationships using deep mutational scanning data sets.

## Installation

Install using **devtools** package:

```{r, eval = F}
# install.packages("devtools")
devtools::install_github("")
```

## Input
- file paths for the two txtfiles containing mutations in labeled (post-selected) and unlabeled (pre-selected) sets (path_l, path_u)
- Prevalence of positive examples in the unlabeled set (py1)

---

## Examples

### Model fitting with a full data set

Here we demonstrate basic usage of the package using sample labeled and unlabeled sequences (10,000 randomly sampled sequences from casp3 protein data sets)

Input:
- path_l, path_u 
- WT (reference sequence)
- py1 

Here we use the paths where the sample txt files are saved. Please replace them with the actual file paths (e.g. `"system.file("extdata", "sample_positive_mutations.txt", package = "pudms",mustWork = T)"` -> `"your dir/filename.txt"`).

```{r}
library(pudms)

# Inputs

## paths for the example txt files
## replace them with paths for your txt files
path_l = system.file("extdata", "sample_positive_mutations.txt", 
                     package = "pudms",mustWork = T)
path_u = system.file("extdata", "sample_unlabeled_mutations.txt", 
                     package = "pudms",mustWork = T)

## WT (reference) sequence
WT = 'MSGISLDNSYKMDYPEMGLCIIINNKNFHKSTGMTSRSGTDVDAANLRETFRNLKYEVRNKNDLTREEIVELMRDVSKEDHSKRSSFVCVLLSHGEEGIIFGTNGPVDLKKITNFFRGDRCRSLTGKPKLFIIQACRGTELDCGIETDSGVDDDMACHKIPVEADFLYAYSTAPGYYSWRNSKDGSWFIQSLCAMLKQYADKLEFMHILTRVNRKVATEFESFSFDATFHAKKQIPCIVSMLTKELYFYHLEHHHHHH*'

## py1 
protein_py1 = 0.2 # change this to the prevalence of positive examples in your unlabeled set
```

1. We generate a protein dataset from the specified paths:

```{r}
smpl_dat = create_protein_dat(path_l = path_l, path_u = path_u, WT = WT)
```

2. Now, we fit a model:

```{r}
# we filter sequences containing mutations with total number < nobs_thresh
fit1 = pudms(protein_dat = smpl_dat,py1 = protein_py1,nobs_thresh = 10) 
```

3. Done!

To view the result in R:
```{r,eval=F}
View(fit1$result_table)
```

To extract the result into the excel file:
```{r,eval=F}
require(xlsx)
write.xlsx(x = fit1$result_table, sheetName = "result", 
           file = "your path/result_table.xlsx")
```


### Model fitting with training and test sets

Here, we will use 90\% of the data to train a model and obtain an ROC curve with remaining 10\% of the data.

1. We split previously generated `smpl_dat` into 90/10 subsamples:

```{r}
cv_smpl = cv_grouped_dat(grouped_dat = smpl_dat,
                         test_idx = 1,
                         nfolds = 10,
                         seed = 1) # seed for reproducibility

tr_smpl = cv_smpl$train_grouped_dat # training set
tt_smpl = cv_smpl$test_grouped_dat  # test set
```

2. We fit a model using a training set `tr_smpl`:

```{r}
cv.r = pudms(protein_dat = tr_smpl,py1 = protein_py1,nobs_thresh = 10)
```

3. We obtain a PU-corrected ROC curve using a test set `tt_smpl`:

``````{r,fig.height=4.5,fig.width=4.5}
roc = corrected_roc_curve(coef = coef(cv.r), 
                          test_grouped_dat = tt_smpl,
                          py1 = protein_py1)
roc$rocplot
```

To save a plot:

```{r, eval = F}
ggsave(filename = "roc.png", plot = roc$rocplot, path = "your path to save")
```


